Visualizations for the manuscript

### 2022-05-10 metanalysis_visualization/manuscript_Figures.Rmd
table 1 add percentages and 95 CI

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly")
paket(listOfPackages)

@@

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
#history_start_date <- as.Date("2019-01-01")
#history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-02-01")
end_date_plots <- as.Date("2021-04-30")
```

@@

```{r}
outputDir <- "../../output/latest_output//"
outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
country_map <- read.delim( file = "../public-data/country_mapping.txt")
```

@@

```{r Extracting age information}
age_table1_all <- table1_all %>% 
  filter(name == "Age" | (name == "Total" & Category == "Sex")) %>% 
  select(name, siteid, before_pandemic, during_pandemic) %>% 
  pivot_longer(cols = c(before_pandemic, during_pandemic), 
               names_to = "period",
               values_to = "values") %>% 
  pivot_wider(names_from = name, 
              values_from = values, 
              id_cols = c(siteid, period)) %>% 
  mutate(mean_age = as.numeric(sub("\\s\\(.*", "", x = Age, perl = TRUE)),
         sd_age = as.numeric(str_extract(pattern = "(?<=\\(sd:\\s)[0-9]{1,2}\\.[0-9]{1,2}(?=\\))", 
                                         string = Age)),
         Total = as.numeric(sub("\\s\\(.*", "", x = Total, perl = TRUE))
  ) %>% 
  group_by(period) %>% 
  summarise(mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
            sd_age = round(sum(Total * sd_age) / sum(Total), 2)) %>% 
  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")")) %>% 
  select(period, Age) %>% 
  pivot_wider(names_from = period, values_from = Age) %>% 
  mutate(Category = "Age", 
         name = "mean (sd)") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

```{r}
median_table1_all <- length_hospitalisation_values_all %>% 
  filter(siteid != "CCHMC") %>% # temporary
  group_by(period) %>% 
  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
  pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
  mutate(Category = "Length of stays", 
         name = "median") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

```{r}
sex_table1_all <- table1_all %>%
  filter( Category == "Sex" & name != "Total") %>%
  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
  group_by( name ) %>%
  summarise( before_pandemic_count = sum( as.numeric( before_counts )), 
             during_pandemic_count = sum( as.numeric( during_counts ))) %>%
  mutate( Category = "Sex") %>%
  select( Category, name, before_pandemic_count, during_pandemic_count )

sex_table1_all <- rbind( sex_table1_all, c("Total", "",sum( sex_table1_all$before_pandemic_count), sum( sex_table1_all$during_pandemic_count)))

dg_table1_all <- table1_all %>%
  filter( Category == "Disorder groups" & name != "Total") %>%
  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
  group_by( name ) %>%
  summarise( before_pandemic_count = sum( as.numeric( before_counts ), na.rm = TRUE), 
             during_pandemic_count = sum( as.numeric( during_counts ), na.rm = TRUE)) %>%
  mutate( Category = "Disorder groups") %>%
  select( Category, name, before_pandemic_count, during_pandemic_count )

table1 <- rbind( sex_table1_all, age_table1_all, median_table1_all, dg_table1_all )

>>>>

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels")
paket(listOfPackages)

@@

```{r}
time_period <-"months" #possible values: days, weeks, months
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-02-01")
end_date_plots <- as.Date("2021-04-30")
```

@@

```{r}
outputDir <- "../../output/latest_output/"
outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
country_map <- read.delim( file = "../public-data/country_mapping.txt")
```

@@

```{r Extracting age information}
age_table1_all <- table1_all %>% 
  filter(name == "Age" | (name == "Total" & Category == "Sex")) %>% 
  select(name, siteid, before_pandemic, during_pandemic) %>% 
  pivot_longer(cols = c(before_pandemic, during_pandemic), 
               names_to = "period",
               values_to = "values") %>% 
  pivot_wider(names_from = name, 
              values_from = values, 
              id_cols = c(siteid, period)) %>% 
  mutate(mean_age = as.numeric(sub("\\s\\(.*", "", x = Age, perl = TRUE)),
         sd_age = as.numeric(str_extract(pattern = "(?<=\\(sd:\\s)[0-9]{1,2}\\.[0-9]{1,2}(?=\\))", 
                                         string = Age)), 
         Total = as.numeric(sub("\\s\\(.*", "", x = Total, perl = TRUE))
  ) %>% 
  group_by(period) %>% 
  summarise(conf_interval = gmodels::ci( mean_age ),
            mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
            sd_age = round(sum(Total * sd_age) / sum(Total), 2), 
            ci_age = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>% 
  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")", as.character(ci_age))) %>% 
  select(period, Age) %>% 
  unique() %>%
  pivot_wider(names_from = period, values_from = Age) %>% 
  mutate(Category = "Age", 
         name = "mean (sd), 95% CI") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

```{r}
median_table1_all <- length_hospitalisation_values_all %>% 
  group_by(period) %>% 
  summarise(median_hospitalisation = mean(len_hospitalisation)) %>% 
  pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
  mutate(Category = "Length of stays", 
         name = "median") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
median_table1_all$before_pandemic_count <- as.character(median_table1_all$before_pandemic_count)
median_table1_all$during_pandemic_count <- as.character(median_table1_all$during_pandemic_count)

mean_table1_all <- length_hospitalisation_values_all %>% 
  group_by(period) %>% 
  summarise(mean_hospitalisation = round( mean(len_hospitalisation), 2), 
            conf_interval = gmodels::ci( len_hospitalisation ), 
            ci_length = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>%
  mutate(mean_hospitalisation = paste0(as.character(mean_hospitalisation), as.character(ci_length))) %>% 
  select(period, mean_hospitalisation) %>%
  unique() %>%
  pivot_wider(names_from = period, values_from = mean_hospitalisation) %>% 
  mutate(Category = "Length of stays", 
         name = "mean [95% CI]") %>% 
  select(Category, name, before_pandemic, during_pandemic) %>% 
  rename(before_pandemic_count = before_pandemic, 
         during_pandemic_count = during_pandemic)
```

```{r}
sex_table1_all <- table1_all %>%
  filter( Category == "Sex" & name != "Total") %>%
  mutate( before_counts = as.numeric(sapply(strsplit( before_pandemic, " "), '[', 1)), 
          during_counts = as.numeric(sapply(strsplit( during_pandemic, " "), '[', 1)), 
          before_total = sum( before_counts ), 
          during_total = sum( during_counts)) %>%
  group_by( name ) %>%
  summarise( before_pandemic_count = sum( as.numeric( before_counts )), 
             during_pandemic_count = sum( as.numeric( during_counts )), 
             before_pandemic_perc  = round(100*before_pandemic_count/before_total, 1), 
             during_pandemic_perc  = round(100*during_pandemic_count/during_total, 1)) %>%
  mutate( Category = "Sex", 
          before_pandemic_count = paste0( before_pandemic_count, " (", before_pandemic_perc, ")"), 
          during_pandemic_count = paste0( during_pandemic_count, " (", during_pandemic_perc, ")")) %>%
  select( Category, name, before_pandemic_count, during_pandemic_count ) %>%
  unique()

dg_table1_all <- table1_all %>%
  filter( Category == "Disorder groups" & name != "Total") %>%
  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
  group_by( name ) %>%
  summarise( before_pandemic_count = sum( as.numeric( before_counts ), na.rm = TRUE), 
             during_pandemic_count = sum( as.numeric( during_counts ), na.rm = TRUE), 
              before_pandemic_perc  = round(100*before_pandemic_count/9696, 1), 
             during_pandemic_perc  = round(100*during_pandemic_count/10561, 1)) %>%
  mutate( Category = "Disorder groups", 
           before_pandemic_count = paste0( before_pandemic_count, " (", before_pandemic_perc, ")"), 
          during_pandemic_count = paste0( during_pandemic_count, " (", during_pandemic_perc, ")"))%>%
  select( Category, name, before_pandemic_count, during_pandemic_count )

table1 <- rbind( sex_table1_all, age_table1_all, median_table1_all, mean_table1_all, dg_table1_all )
